### 2022-01-12 paper/paper.tex
methodology: expound hand tracking

A computer vision algorithm will be chosen as a starting point for detecting,
and tracking fingers within a video.

The first component this algorithm will have is the ability to detect fingers
within a video. This would involve capturing exactly where the finger is in the
current frame. This data would then be used for the next component of the
algorithm: tracking. This component would track the movement of the finger as it
performs a typing motion. This algorithm should have the ability to separate the
different fingers and track the movement of the ten fingers individually.

The specific algorithm used for finger detection and tracking will be the
solution offered by MediaPipe. More details about the specific algoritihm can be
found in Section~\ref{section:rrl-mediapipe}.

@@

\subsection{Create typing test sequences}
Typing test sequences are strings that will be used in testing the user in their
ability to type. The test sequences that will be used in the trainer will come
from text found in the public domain obtained from Project Gutenberg and the
Internet Archive. Sentences from these text will be isolated and used as the
test sequences. There will be a total of 20 test sequences that will be
gathered. An example test sequence is ``What of it, if some old hunks of a
sea-captain orders me to get a broom and sweep down the decks?'' from Moby Dick
by \cite{moby-dick}. The 19 other test sequences can be found in the appendix.

>>>>

A computer vision algorithm will be chosen as a starting point for detecting,
and tracking fingers within a video. The specific algorithm used for finger
detection and tracking will be the solution offered by MediaPipe dubbed
MediaPipe Hands. This algorithm is composed of multiple ML models working in
conjunction to be able to detect the different parts of the hands and track them
accurately.

\subsubsection{Palm Detection Model}
The first model, the Palm Detection Model detects the initial hand locations
using a single shot detector model based on the paper by \cite{ssd}. This
model achieves an average precision of 95.7\% in palm detection.

MediaPipe Hands detects the palms first, instead of the whole hands with one
model because the hands lack high contrast patterns. This prevents the model to
detect the hand accurately. In addition, detecting a palm is simpler compared to
detecting hands with articulated fingers since estimating a bounding box around
a rigid objects, i.e.\ a palm, is much simpler. Furthermore, a palm can be
modeled using only square anchors reducing the number of anchors by a factor of
3--5.

\subsubsection{Hand Landmark Model}
After the palms have been detected and an appropriate anchor has been
established, the Hand Landmark Model pinpoints 21 3D hand-knuckle coordinates
inside the detected hand. This is done using direct coordinate prediction.

This model was trained using 30,000 manually annotated, real-world images with
21 3D hand-knuckle coordinates. Using this information, the model can also
accurately add landmarks to partially visible hands and hands with
self-occlusion. This is also made possible by the model's consistent internal hand
pose representation.

@@

\subsection{Typing Test Sequences}
Typing test sequences are strings that will be used in testing the user in their
ability to type. The test sequences that will be used in the trainer will come
from text found in the public domain obtained from Project Gutenberg and the
Internet Archive. Sentences will be isolated from these text and used as test
sequences if they fit a criteria.

The criteria for choosing test sequences is as follows: (1) $\approx80\%$ of the
characters in the keyboard is present in a test sequence. (2) The number of
words in a test sequence do not exceed 25. (3) Numbers and punctuations should
be present in at least $\approx30\%$ of the total test sequences.

There will be a total of 10 test sequences that will be gathered. An example
test sequence is ``What of it, if some old hunks of a sea-captain orders me to
get a broom and sweep down the decks?'' from Moby Dick by \cite{moby-dick}. The
9 other test sequences can be found in the appendix.
